Example case study in GMSE

### 2018-08-20 notebook/ms/SI3.Rmd
Address Reviewer 1 minor comment 5 with a new section of SI3 on non-default parameter values

Simulations using the default GMSE sub-models described above are run using the `gmse` function, which offers a range of options for setting parameter values (see Table 1 for some select examples). Output of `gmse` is an exhaustive list that includes all resources and observations, all stakeholder decisions and actions, and all landscape properties in each time step of the simulation (see [SI7](SI7.pdf) for a description of key data structures). Results are most easily interpreted visually, so a summary of simulation dynamics is plotted by default (the plot can also be called using the `plot_gmse_results` function, and summaries of results can be obtained using `gmse_summary` and `gmse_table`). An example below shows how simulations are set and interpreted.

@@

|  res_move_type  |  1      | Type of resource movement (default is up to res_movement cells in any direction)                 |
|  observe_type   |  0      | Type of resource observation (default is density-based; i.e, counting a subset on the landscape) |

@@

Table 1: Select parameter values for initialising generalised management strategy evaluation simulations. See [below](#nondefaults) for explanation of non-default values.

@@

Non-default options in default GMSE sub-models {#nondefault}
--------------------------------------------------------------------------------

Within the default submodels of GMSE, there are several non-default options that can be set using arguments to `gmse` and `gmse_apply`. Brief explanations of these non-default options appear in the [GMSE documentation](https://cran.r-project.org/web/packages/GMSE/GMSE.pdf) (also [listed](https://confoobio.github.io/gmse/reference/gmse.html) on the [GMSE website](https://confoobio.github.io/gmse/)). Here we further explain some of the less obvious or less intuitive options available in GMSE v0.4.0.6, for which additional explanation beyond the GMSE documentation may be warranted. Future versions of GMSE are expected to expand upon these options.

**res_move_type**: Resources in default GMSE v0.4.0.7 sub-models can move in one of three ways. These different movement types are listed below, and are taken as arguments to `gmse` (e.g., `gmse(res_move_type = 2)`).

  0. *No movement*. Given this option, resources do not move at all in the default resource sub-model, and instead remain on the location at which they are initialised.
  1. *Uniform movement in any direction up to `res_movement` cells away from a resource's original location* (default). In uniform movement, the resource moves to a new cell within some distance `res_movement` of its original location in each time step. Distance therefore includes diagonal movement, such that a resource capable of moving `res_movement = 1` cell away can go to any of the eight adjacent cells around its current location; it can also remain on its current cell. Similarly, a resource capable of moving `res_movement = 2` cells away can go to any of the 24 cells surrounding its current location, or remain on its current cell. The distance (in cells) moved in the x-direction (i.e., left to right) and y-direction (i.e., up and down), is sampled from a uniform distribution, meaning that all reachable cells (including the resource's current cell) have equal probability of being selected as the resource's new cell location.
  2. *Poisson selected movement distance and direction*. In this movement type, the distance (in cells) moved in the x-direction and the y-direction are sampled from a Poisson distribution using `res_movement` as the rate parameter (i.e., the mean movement distance in each direction). Using this movement type is not generally recommended because it is not likely to be biologically realistic (e.g., diagonal movement is rare because selection of both high x-direction and high y-direction values is unlikely).
  3. *Uniform movement in any direction up to `res_movement` cells away from a resource's original location `res_movement` times.* Like option (1), when moving, resources can travel in any direction up to `res_movement` cells from their current location with equal probability. What is different about this option is that uniform movement in any direction is occurs $C =$ Poisson(`res_movement`) times. In other words, when a resource moves, it first decides how many different cells it will visit ($C$), as determined by sampling from a Poisson distribution with a rate parameter of `res_movement` (i.e., the mean number of visited cells will be `res_movement`). Upon leaving a cell, the resource will move in the same way as defined in option (1) above, by choosing a cell within `res_movement` distance (in cells) of its current cell. It continues moving in this way until $C$ cells have been visited. This type of movement has been used in some individual-based models of plant-pollinator interactions [e.g., @Duthie2013; @Wilson2003].

**res_death_type**: Resources can be removed from the model in one of three ways within the default resource sub-model, thereby simulating resource death. It is important to note that `res_death_type` refers only to removal that occurs within the resource sub-model; removal from user actions (e.g., death from shooting) occurs within the user sub-model.

  1. *Density-independent removal*. Given this option, there is no density-dependent removal of resources within the resource sub-model. Resources are simply removed with a probability of `removal_pr` each time the submodel is run. This option should be used with great care because in the absence of density-dependence, it is possible for the population of resources to increase without limit and thereby greatly slow down simulation times. 
  2. *Density-dependent removal*. Given this option, removal of resources within the resource sub-model is caused entirely by density effects. The probability that a resource is removed is a function of the resource carrying capacity set using the `res_death_K` parameter. If the number of resources ($N$) is greater than `res_death_K` ($K$), then the probability that an individual resource is removed is defined by $(N - K) / N$. Removal then occurs independently for each resource based on this probability.
  3. *Density-independent and density-dependent removal*. This option allows a combination of both options (1) and (2), which affect removal independently. In other words, each resource is assigned a probability of removal `removal_pr` as in (1), and then independently assigned another probability of removal based on `res_death_K` as in (2). If removal occurs as a consequence of (1), (2), or both, then the resource is removed.

**observe_type**: Observation of resources in the observation model can occur in one of four ways.

**obs_move_type**:

>>>>

Simulations using the default GMSE sub-models described above are run using the `gmse` function, which offers a range of options for setting parameter values (see Table 1 for some select examples). Output of `gmse` is an exhaustive list that includes all resources and observations, all stakeholder decisions and actions, and all landscape properties in each time step of the simulation (see [SI7](SI7.pdf) for a description of key data structures).

@@

|  res_death_type |  1      | Rules affecting resource death (default is density-dependent; [see below](#nondefault))          |
|  res_move_type  |  1      | Type of resource movement (default moves uniformly in any direction; [see below](#nondefault))   |
|  observe_type   |  0      | Type of resource observation (default is density-based; [see below](#nondefault))                |
|  obs_move_type  |  1      | How agents (manager and stakeholders) move (typically ignored; [see below](#nondefault))         |

@@

Table 1: Select parameter values for initialising generalised management strategy evaluation simulations. See [below](#nondefault) for explanation of non-default values of resource death and resource observation processes.

Results are most easily interpreted visually, so a summary of simulation dynamics is plotted by default (the plot can also be called using the `plot_gmse_results` function, and summaries of results can be obtained using `gmse_summary` and `gmse_table`). An example below shows how simulations are set and interpreted.

@@

Default and non-default options in default GMSE sub-models {#nondefault}
--------------------------------------------------------------------------------

Within the default sub-models of GMSE, there are several non-default options that can be set using arguments to `gmse` and `gmse_apply`. Brief explanations of these non-default options appear in the [GMSE documentation](https://cran.r-project.org/web/packages/GMSE/GMSE.pdf) (also [listed](https://confoobio.github.io/gmse/reference/gmse.html) on the [GMSE website](https://confoobio.github.io/gmse/)). Here we further explain some of the less obvious or less intuitive options available in GMSE v0.4.0.7, for which additional explanation beyond the GMSE documentation may be warranted for application to case studies such as the Taiga Bean Geese case study example demonstrated above. Future versions of GMSE are expected to expand upon these options.

**res_move_type**: Resources in default GMSE v0.4.0.7 sub-models can move on the landscape in one of three ways. These different movement types are listed below, and are taken as arguments to `gmse` (e.g., `gmse(res_move_type = 2)`).

  0. *No movement*. Given this option, resources do not move at all in the resource sub-model, and instead remain on the location at which they are initialised.
  1. *Uniform movement in any direction up to `res_movement` cells away from a resource's original location* (default). In uniform movement, the resource moves to a new cell within some distance `res_movement` of its original location in each time step. Distance therefore includes diagonal movement, such that a resource capable of moving `res_movement = 1` cell away can go to any of the eight adjacent cells around its current location; it can also remain on its current cell. Similarly, a resource capable of moving `res_movement = 2` cells away can go to any of the 24 cells surrounding its current location, or remain on its current cell. The distance (in cells) moved in the x-direction (i.e., left to right) and y-direction (i.e., up and down), is sampled from a uniform distribution, meaning that all reachable cells (including the resource's current cell) have equal probability of being selected as the resource's new cell location.
  2. *Poisson selected movement distance and direction*. In this movement type, the distance (in cells) moved in the x-direction and the y-direction are sampled from a Poisson distribution using `res_movement` as the rate parameter (i.e., the mean movement distance in each direction). Using this movement type is not generally recommended because it is not likely to be biologically realistic (e.g., diagonal movement is rare because selection of both high x-direction and high y-direction values is unlikely).
  3. *Uniform movement in any direction up to `res_movement` cells away from a resource's original location `res_movement` times.* Like option (1), when moving, resources can travel in any direction up to `res_movement` cells from their current location with equal probability. What is different about this option is that uniform movement in any direction occurs $C =$ Poisson(`res_movement`) times. In other words, when a resource moves, it first decides how many different cells it will visit ($C$), as determined by sampling from a Poisson distribution with a rate parameter of `res_movement` (i.e., the mean number of visited cells will be `res_movement`). Upon leaving a cell, the resource will move in the same way as defined in option (1) above, by choosing a cell within `res_movement` distance (in cells) of its current cell. It continues moving in this way until $C$ cells have been visited. This type of movement has been used in some individual-based models of plant-pollinator interactions [e.g., @Duthie2013; @Wilson2003].

**res_death_type**: Resources in GMSE v0.4.0.7 can be removed from the model in one of three ways within the default resource sub-model, thereby simulating resource death. It is important to note that `res_death_type` refers only to removal that occurs within the resource sub-model; removal from user actions (e.g., death from shooting) occurs within the user sub-model.

  1. *Density-independent removal*. Given this option, there is no density-dependent removal of resources within the resource sub-model. Resources are simply removed with a probability of `removal_pr` each time the sub-model is run. This option should be used with great care because in the absence of density-dependence, it is possible for the population of resources to increase without limit and thereby greatly slow down simulation times. 
  2. *Density-dependent removal* (default). Given this option, removal of resources within the resource sub-model is caused entirely by density effects. The probability that a resource is removed is a function of the resource carrying capacity set using the `res_death_K` parameter. If the number of resources ($N$) is greater than `res_death_K` ($K$), then the probability that an individual resource is removed is defined by $(N - K) / N$. Removal then occurs independently for each resource based on this probability.
  3. *Density-independent and density-dependent removal*. This option allows a combination of both options (1) and (2), which affect removal independently. In other words, each resource is assigned a probability of removal `removal_pr` as in (1), and then independently assigned another probability of removal based on `res_death_K` as in (2). If removal occurs as a consequence of (1), (2), or both, then the resource is removed.

**observe_type**: Observation of resources in default GMSE v0.4.0.7 sub-models can occur in one of four ways. Each observation type mimics some process of resource observation, with potential error affected by observation intensity.

  0. *Density-based observation*. In this option, managers observe resources on a subset of the landscape; subset size is determined by the manager's view as set using the `agent_view` parameter. In practice, this is done by having the manager pick a cell on the landscape at random, then count all of the resources within a distance (in any direction) of `agent_view` cells. Managers then extrapolate the density of resources within this subset to estimate the total number of resources on the landscape. For example, if the manager has an `agent_view` of 1, then they are capable of seeing all of the resources on nine landscape cells (the cell on which they are located, and each adjacent cell). If they count that there are 90 resource on these nine cells, then they calculate a resource density of 10 per cell. They then multiply the density of 10 by the total number of landscape cells to estimate the number of resources on the landscape. The error of this estimate naturally decreases with increasing `agent_view`, such that when the manager is capable of viewing the entire landscape, observation error is zero. If desired, the `times_observe` (default equals 1) option can also be used to allow managers to sample more than one landscape subset in their estimate (e.g., if `times_observe = 2`, then density is estimated from two different locations, with the mean of individual estimates used for a combined estimate of population density). 
  1. *Mark-recapture observation*. This option simulates the process of a mark-recapture analysis. In this process, managers randomly sample `fixed_mark` resources in the population; sampling occurs without replacement and without any spatial bias, and if `fixed_mark` is greater than the total number of resources on the landscape, then managers sample all resources. The manager then randomly samples `fixed_recapt` resources, again without replacement or spatial bias. The sampled resources from `fixed_mark` and `fixed_recapt` are interpreted as marked and recaptured resources, respectively, and a Chapman estimate [@Pollock2016] is then used in the manager model to calculate estimated population size form these observation data. Error in this estimate can be naturally minimised if both `fixed_mark` and `fixed_recapt` are very high.
  2. *Sample linear transect*. This option simulates sampling from a linear transect of the landscape. The manager samples an entire set of cell rows on the landscape and counts the total resources on all cells; the manager then continues onto the next set of landscape rows until the entire landscape has been sampled. The number of rows observed in each sample is defined by `agent_view`. For example, if the landscape is of the dimensions $100 \times 100$ cells, and `agent_view = 2`, then the manager will first observe all resources in the upper 200 cells (first $2 \times 100$ cells in the upper two landscape cell rows), then move onto the next 200 cells, repeating the process 50 times until all rows have been sampled. This process of sampling will only generate observation error if `res_move_obs = TRUE`, which it is by default. In this case, resources can move on the landscape (according to the rules of `res_move_type` and distance of `res_movement`) in between transect observation, potentially causing some resources to escape observation and others to be observed more than once. 
  3. *Sample block transect*. This option is identical to the *sample linear transect* option above, except that instead of sampling an entire row of a landscape, the manager samples one square block of the landscape at a time. The cell length of each block side equals `agent_view`, meaning that the manager can observe all of the resources on `agent_view * agent_view` cells at a time. The manager proceeds to observe the entire landscape, block by block, until all cells on the landscape have been covered. For example, if `agent_view = 25`, and the landscape is of the dimensions $100 \times 100$, then the manager will sample four total blocks of the landscape, one at a time, and use the total number of resources observed as the estimate of resource density on the landscape. As with linear transect sampling, if `res_move_obs = TRUE` (which it is by default), then resources can move on the landscape between block samples, potentially causing observation error if some resources are missed due to movement, or counted multiple times.

**obs_move_type**: Observation of agents (manager and users) in default GMSE v0.4.0.7 sub-models can occur in the same four ways as allowed in `res_move_type` (default `obs_move_type = 1`), with movement distance defined by the parameter `agent_move` (default 50). Currently, the only effect of this movement is to allow for spatially autocorrelated sampling of resources by the manager in the observation model. If the manager is using density-based sampling (`observe_type = 0`) and sampling more than one subset of the landscape (`times_observe > 1`), then the manager's movement between subsets is governed by `res_move_type` (movement rules) and `agent_move` (movement distance). Hence, to have the manager observe the population many times in one area of the landscape, the `times_observe` option should be increased above one and `agent_move` should be decreased to a low value.
